---
title: "Exercise Prediction Model Analysis"
author: "Vinay"
date: "October 24, 2015"
output: html_document
---

#Summary 

This project's goal is to build model to predict activity quality from activity monitors. Data provided by http://groupware.les.inf.puc-rio.br/har is used in this poject to train the model. Based on this data training set accuracy was noted as 99%. 

#Prepare Environment
```{r, echo=FALSE}
library(caret)
library(randomForest)
library(doParallel)
cl<-makeCluster(detectCores())
registerDoParallel(cl)
```


#Load and Clean Data
Load data and remove NA and non numberic variables. Exclude variables that have time, or names in it, also new_window. Columns 1..6
```{r}
trainRawData <- read.csv("pml-training.csv",na.strings=c("NA",""))
newdataRawData <- read.csv("pml-testing.csv",na.strings=c("NA",""))
NAs <- apply(trainRawData,2,function(x) {sum(is.na(x))}) 
trainData <- trainRawData[,which(NAs == 0)]
newData <- newdataRawData[,which(NAs == 0)]

# Exclude variables that have time, or names in it, also new_window. Columns 1..6
removeIndex <- as.integer(c(1,2,3,4,5,6))
trainData <- trainData[,-removeIndex]
newData <- newData[,-removeIndex]

```

#Prediction Study Design
I consider sample set as small due to having 19622 obs on 6 persons. So I use cross validation in study design. Approximately 70% of the data is used for the training and remaining 30% for cross validation.

```{r}
trainIndex <- createDataPartition(y = trainData$classe, p=0.7,list=FALSE)
trainSet <- trainData[trainIndex,]
testSet <- trainData[-trainIndex,]
```

#Training
Train model using radom Forecast and cross validation as train control.
```{r}
trainCtrl = trainControl(method = "cv", number = 4)
modelFit <- train(trainSet$classe ~.,data = trainSet, method="rf", trControl = trainCtrl)
print(modelFit, digits=3)

```

#Testing 
now test the model using cross validation data. Confusion matrix and statistics are shown below. Testing model using test data shows accuracy of 99.8% with confidence interval 99.6% to 99.9%. 
```{r}
predictionsTest <- predict(modelFit, testSet)
print(confusionMatrix(predictionsTest, testSet$classe), digits=4)
```

#out of sample error
Previous section showed various statistics when model is used to predict classe on a data that model never saw. Based on those statistics out of sample error is calcualted as (1 - 0.9983) = 0.0017.

#Predict on new Data 
```{r}
predictedclasse <- predict(modelFit, newData)
length(predictedclasse)
```

#Prediction Assignment Submission
This section is for generating files for assignment submission.

```{r}
pml_write_files = function(predictedclasse){
  n = length(predictedclasse)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(predictedclasse[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictedclasse)

```


